---
title: "ESM 244 Lab"
author: "Camille Herrera"
date: "March 7, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}

library(beepr) # beep (on CRAN)
library(datapasta) # copy/paste (on CRAN)
library(tidyverse) 
library(igraph) # network stuff (on CRAN)
library(multiplex) # network stuff (on CRAN)
library(ggraph) # More graph stuff (on CRAN)
library(gt) # tables (get development version!!!)
# remotes::install_github("rstudio/gt")
library(ggalluvial) # Sankeys (on CRAN)
library(readxl) # Get .xls (on CRAN)
library(praise) # Get/share praise! 

# Note: may need to update dependencies (I usually choose yes to all...)


```

###Part 1. Network analysis example (Les Miserables character connections)

Get the data
```{r}

# Stored as a .gml file (graph modeling language) format supports network data
# Use multiplex::read.gml()

lm_df <- read.gml("lesmis.gml")

# Convert to graph (igraph) with nodes & edges
les_mis <- graph_from_data_frame(lm_df, directed = FALSE) # means no direction in graph theory we are using
beep(sound = 8) # ?beep to see options but don't pipe into the beep sound, otherwise it will store i as an audio object
praise()

```

Find some quantiative metrics of graph (as in graph theory):

```{r}
#graph diameter:
diameter(les_mis)

farthest_vertices(les_mis)

```

Plot:
```{r}

plot(les_mis)

plot(les_mis, vertex.color = "orange",
     vertex.size = 5,
     vertex.label.cex = 0.5)

```

A little more customized graphs using gg formats:
```{r}

ggraph(les_mis, layout = "kk") + #equivalent of ggplot = 
  geom_edge_link()

ggraph(les_mis, layout = "kk") + #equivalent of ggplot = 
  geom_edge_link() +
  geom_node_text(aes(label = name), size = 2, color = "white") +
  theme_dark()

ggraph(les_mis, layout = "linear") +
  geom_edge_arc(alpha = 0.8) +
  geom_node_text(aes(label = name), angle = 90, size = 2, hjust = 1) +
  theme_graph()

ggraph(les_mis, layout = "tree") +
  geom_edge_fan(color = "red") +
  geom_node_text(aes(label = name), size = 2, color = "blue", angle = 45) +
  theme_void()
beep(3)

```

Or the ones that we're most used to seeing:
```{r}

ggraph(les_mis, layout = "linear", circular = TRUE) +
  geom_edge_arc(alpha = 0.5)

ggraph(les_mis, layout = "linear", circular = TRUE) +
  geom_edge_arc(alpha = 0.5) +
  geom_node_point(aes(color = name), show.legend = FALSE) +
  geom_node_text(aes(label = name), size = 3,  hjust = "outward") +
  theme_void()

# can show hieracharial clusters using graph theory
# can be these are two variables that have a correction greater than 0.8, can use what we did with matrices in an earlier lab

```

###Part 2. Sankey diagrams

- alluvial #similar syntax to ggplot
- ggalluvial
- NetworkD3
- ggforce

```{r}

sdf <- read_csv("sankey_df.csv") %>% 
  select(-X1)

```

```{r}

ggplot(sdf, aes(y = weight, axis1 = before, axis2 = after)) +
  geom_alluvium(aes(fill = before, color = before), show.legend = FALSE)

ggplot(sdf, aes(y = weight, axis1 = before, axis2 = after)) +
  geom_alluvium(aes(fill = before, color = before), show.legend = FALSE) +
  geom_stratum(width = 1/5, color = "gray") + #what proportion of the graph will this be in
  geom_text(stat = "stratum", label.strata = TRUE) 

ggplot(sdf, aes(y = weight, axis1 = before, axis2 = after)) +
  geom_alluvium(aes(fill = before, color = before), show.legend = FALSE) +
  geom_stratum(width = 1/5, color = "gray") + 
  geom_text(stat = "stratum", label.strata = TRUE) +
  scale_fill_manual(values = c("purple", "blue", "green")) +
  scale_color_manual(values = c("purple", "blue", "green"))

ggplot(sdf, aes(y = weight, axis1 = before, axis2 = after)) +
  geom_alluvium(aes(fill = before, color = before), show.legend = FALSE) +
  geom_stratum(width = 1/5, color = "gray") + 
  geom_text(stat = "stratum", label.strata = TRUE) +
  scale_fill_manual(values = c("purple", "blue", "green")) +
  scale_color_manual(values = c("purple", "blue", "green")) +
  scale_x_discrete(limits = c("Before", "After"), expand = c(0,0)) + # how to get rid of areas below 0,0 axes
  theme_void()

```

###Part 3. Creating your own tibbles

tibbles are dfs that have some extra functionality compared to what used to be called dataframes
tibble and df used interchangably in real world
good to have tiny ddf to try code on
```{r}

my_tibble <- tribble(
  ~camille, ~made, ~this, ~awesomething,
  1, "yes", 0, 10000,
  2, "no", 4, 50000,
  3, "maybe", 10, 40000,
  4, "yes", 9, 15000
)
beep(2)

# can recreate a reproducible example

```

Check out Jenny Bryan's awesome package 'reprex' - incentivizes people to make little stand-alone code that people are having problems with can run and share to get help (like overstack or R Studio Community)

###3b. datapasta

Can I copy and paste data from outside sourcees and have R store it as a df or tibble?


```{r}

weather_data <- tibble::tribble(
                                                     ~X,          ~Location, ~Min, ~Max,
                                       "Partly cloudy.",         "Brisbane",   19,   29,
                                       "Partly cloudy.", "Brisbane Airport",   18,   27,
                                     "Possible shower.",       "Beaudesert",   15,   30,
                                       "Partly cloudy.",        "Chermside",   17,   29,
                       "Shower or two. Possible storm.",           "Gatton",   15,   32,
                                     "Possible shower.",          "Ipswich",   15,   30,
                                       "Partly cloudy.",    "Logan Central",   18,   29,
                                        "Mostly sunny.",            "Manly",   20,   26,
                                       "Partly cloudy.",    "Mount Gravatt",   17,   28,
                                     "Possible shower.",            "Oxley",   17,   30,
                                       "Partly cloudy.",        "Redcliffe",   19,   27
                       )

weather_data <- rename(weather_data, Condition = X)

# create shortcuts > Tools > Addins > Browse Addins > Keyboard shortcuts
camille_tries_againtibble::tribble(
                                     ~X,          ~Location, ~Min, ~Max,
                       "Partly cloudy.",         "Brisbane",   19,   29,
                       "Partly cloudy.", "Brisbane Airport",   18,   27,
                     "Possible shower.",       "Beaudesert",   15,   30,
                       "Partly cloudy.",        "Chermside",   17,   29
                     )


write_csv(camille_datapasta, "camille_storing_things.csv")
# want to create a readme file with metadata

```

###3c. Beautiful customized tables with 'gt'

```{r}

weather_data %>% 
  gt() 

weather_data %>% 
  gt() %>% 
  tab_header(
    title = "Camille's Table",
    subtitle = "Cool subtitle"
  )

weather_data %>% 
  gt() %>% 
  tab_header(
    title = "Camille's Table",
    subtitle = "Cool subtitle"
  ) %>% 
  fmt_number(
    columns = vars(Min),
    decimals = 2
  ) %>% 
  cols_move_to_start(
    columns = vars(Location)
  )

#not the most concise code but most digestible and easy to customize

weather_data %>% 
  gt() %>% 
  tab_header(
    title = "Camille's Table",
    subtitle = "Cool subtitle"
  ) %>% 
  fmt_number(
    columns = vars(Min),
    decimals = 2
  ) %>% 
  cols_move_to_start(
    columns = vars(Location)
  ) %>% 
  data_color(
    columns = vars(Min),
    colors = scales::col_numeric(
      palette = c(
        "darkblue", "blue"
      ), 
      domain = c(14,20)
    )
  )

#changes table into heat mat

weather_data %>% 
  gt() %>% 
  tab_header(
    title = "Camille's Table",
    subtitle = "Cool subtitle"
  ) %>% 
  fmt_number(
    columns = vars(Min),
    decimals = 2
  ) %>% 
  cols_move_to_start(
    columns = vars(Location)
  ) %>% 
  data_color(
    columns = vars(Min),
    colors = scales::col_numeric(
      palette = c(
        "darkblue", "blue"
      ), 
      domain = c(14,20)
    )
  ) %>% 
  tab_options(
    table.background.color = "purple", 
    heading.background.color = "black",
    column_labels.background.color = "gray"
  )

```

###3d. Reading in directly from a url

make sure you write it in with history of when you accessed it

```{r}
nuclear <- read_delim("https://www.nrc.gov/reading-rm/doc-collections/event-status/reactor-status/PowerReactorStatusForLast365Days.txt", delim = "|", col_names = TRUE)

# Because you always want a version of what you were working with, I recommend writing that to a csv (or a text file, or whatever)

# write_csv(nuclear, "nuclear.csv")

```

```{r}

pesticides <- read_xls("PesticideResidues.xls")

pest2 <- read_xls("PesticideResidues.xls", skip = 2, col_names = TRUE)

pest <- pest2 %>% 
  janitor::clean_names()

### Do some wrangling:

crops <- pest %>%
  filter(commodity =="KALE", 
         !is.na(grower_city)) %>% 
  separate(grower_city, 
           c("grow_city", "grow_state"), sep = ",") %>% 
  separate(collection_site_city, 
           c("market_city", "market_state", 
             sep = ",")) %>% 
  group_by(organic_commodity, grow_city, market_city) %>% 
  tally()

# Make a Sankey diagram!

ggplot(crops, aes(y = n, axis1 = organic_commodity, axis2 = grow_city, axis3 = market_city)) +
  geom_alluvium(aes(fill = organic_commodity, color = organic_commodity), show.legend = FALSE) +
  geom_stratum() +
  geom_text(stat = "stratum", label.strata = TRUE, size =2)
beep(3)

```

Also packages and function to work with text files and googlesheets

